---
title: 'data 621 assignment #3'
author: "Tyler Baker"
date: "2022-11-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(pscl)
```

# Data 621
## Assignment 3
### Tyler Baker,

## Objective
Your objective is to build a binary logistic regression model on the training data set to predict whether the
neighborhood will be at risk for high crime levels. You will provide classifications and probabilities for the
evaluation data set using your binary logistic regression model.

## Data Exploration
Describe the size and the variables in the crime training data set.

### Load Data
```{r}
crime_test_df <- read.csv("C:\\Users\\tyler\\Downloads\\DATA 621\\crime-training-data_modified.csv")
crime_eval_df <- read.csv("C:\\Users\\tyler\\Downloads\\DATA 621\\crime-evaluation-data_modified.csv")
```

### Basic Exploration
Structure of data frame
```{r}
str(crime_test_df)
```

Basic summary statistics
```{r}
summary(crime_test_df)
```

### Variable Visualizations
We will use jitterplots because they account for overlap, which happen in basic plots.

#### zn
zn is the proportion of residential land zoned for large lots (> 25,000 sq.ft.).

```{r}
graphics::plot(jitter(target, .25) ~ jitter(zn), crime_test_df, xlab="zn", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### indus
The proportion of non-retail business acres per suburb.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(indus), crime_test_df, xlab="indus", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### chas
Does the suburb border the Charles river?

```{r}
graphics::plot(jitter(target, .25) ~ jitter(chas), crime_test_df, xlab="chas", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### nox
nitrogren oxides concentration (parts per 10 million).

```{r}
graphics::plot(jitter(target, .25) ~ jitter(nox), crime_test_df, xlab="nox", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### rm 
average number of rooms per dwelling.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(rm), crime_test_df, xlab="rm", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### age
proportion of owner-occupied units built prior to 1940.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(age), crime_test_df, xlab="age", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### dis
weighted mean of distances to five Boston employment centers.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(dis), crime_test_df, xlab="dis", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### rad
index of accessibility to radial highways.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(rad), crime_test_df, xlab="rad", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### tax
full-value property tax rate per $10,000.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(tax), crime_test_df, xlab="tax", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### ptratio
pupil-teacher ratio by town.

```{r}
graphics::plot(jitter(target, .25) ~ jitter(ptratio), crime_test_df, xlab="ptratio", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```


#### lstat
lower status of the population (percentage)

```{r}
graphics::plot(jitter(target, .25) ~ jitter(lstat), crime_test_df, xlab="lstat", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```

#### medv
median value of owner-occupied homes in $1000s

```{r}
graphics::plot(jitter(target, .25) ~ jitter(medv), crime_test_df, xlab="medv", 
            ylab="target (0 if below median crime rate, 1 if above median crime rate)", pch=".",col='red')
```


## Model Building

I will use a logistical regression and probit regression and compare the results. I will work backwards with each, meaning that I will first include all of the variables and then remove them based on significance codes.

### Logistical Model

```{r}
logit <- glm(target ~ zn + indus + chas + nox + rm + age + dis + rad + tax + ptratio + lstat + medv, data = crime_test_df, family = "binomial")
```

```{r}
summary(logit)
```

By looking at the significance codes we can eliminate indus, chas, rm, and lstat.

```{r}
new_logit <- glm(target ~ zn + nox + age + dis + rad + tax + ptratio + medv, data = crime_test_df, family = "binomial")
```

```{r}
summary(new_logit)
```


### Probit Regression

```{r}
probit <- glm(target ~ zn + indus + chas + nox + rm + age + dis + rad + tax + ptratio + lstat + medv, data = crime_test_df, family = "binomial"(link = "probit"))
```

```{r}
summary(probit)
```
By looking at the significance codes in the summary we can remove zn, indus, chas, rm, lstat.

```{r}
new_probit <- glm(target ~  nox + age + dis + rad + tax + ptratio + medv, data = crime_test_df, family = "binomial"(link = "probit"))
summary(new_probit)
```

### Comparing the models

#### McFadden's Psuedo R^2

```{r}
pR2(new_logit)
```
The logistical regression gave us a McFadden score of 0.69. This means is does a decent job at predicting the target variable.

```{r}
pR2(new_probit)
```
The probit regression returned a McFadden score slightly lower than the logisticall regression. Thus, the logistical regression model has stronger predictive power. We will chose that one to predict.

## Forecasting

Now we will use are model to predict if a neighborhood is above or below the median crime rate.

```{r}
probabilities <- new_logit %>% predict(crime_eval_df, type="response")
predicted.classes <- ifelse(probabilities > 0.5, 1, 0)
```

```{r}
crime_eval_df$predictions <- predicted.classes
```

```{r}
crime_eval_df
```